The Dynamics of Anthropomorphic Effects in Human-Robot Interaction

### 2014-03-19 paper.tex
new intro and overview completed

%{Intro when start with anthropomorphic design}
The Uncanny Valley hypothesis \citep{mori_uncanny_1970} proposes that human-like designs of objects and non-human agents evoke a positive feeling of affinity in the human observer / user, which in turn reflects the experience of emotional engagement with and feelings of empathy for the non-human agent \citep{cheetham_human_2011}. The positive effect of so-called \textbf{anthropomorphic form} is exploited in the design of artificial agents (\eg digital characters, robots) for several reasons. Amongst others, anthropomorphic design should serve to elicit this positive valence as it is predicted in the Uncanny Valley hypothesis, and it should further foster people's acceptance of these artificial agents, and enable effective (human-like and especially human-social) interaction. 
However, at a certain point as the non-human agent approaches but fails to closely resemble a human, the evoked valence in the human user suddenly turns into a negative uncanny feeling of what \cite{mori_uncanny_1970} describes as disgust and revulsion. This, in turn, might lead to rejection of the non-human but human-like agent.
The Uncanny Valley hypothesis predicts that with further increasing the human-likeness of the non-human agent such that it appears so realistic that it is not distinguishable from a human but perceived to be human, the effect is again a positive one, reflecting feelings and engagement that are similar as those evoked by a human-human interaction.
Of course, roboticists and developers try to avoid ``falling'' into this Uncanny Valley with their human-like designs of artificial agents. Consequently, there is a considerable research effort into better understanding and defining the boundaries of anthropomorphic design. However, findings have been inconsistent \citep{cheetham_human_2011}.

A phenomenon that is often studied related to the human-like design of robots is \textbf{anthropomorphism}. Besides a feeling of affinity, and empathy, anthropomorphism can be understood as being one of the resulting (and  more or less desired) effects of the anthropomorphic design of artificial agents (up to the first positive peak). The essence of anthropomorphism is the human user's perception of human-likeness in the non-human agent. Logically, anthropomorphism is said to be enhanced with increased human-like design of the agent. There are various other reasons and explanations for why people tend to anthropomorphize non-human agents. In general, as anthropomorphism emerges in a (real or imagined) interaction \citep{persson_anthropomorphism_2000} between at least two players (the human who anthropomorphizes and the non-human agent which is anthropomorphized) there are also two main perspectives on the phenomenon: an artifact-centered (design) and a human-centered (cognitive, psychological) one. The Uncanny Valley hypothesis can be categorized as being part of first, artifact-centered approach, to understand anthropomorphism.

We observed anthropomorphism in several short- and long-term human-robot interaction studies that we carried out with embodied robots, mostly in real-world settings.\footnote{Note that it was not the main purpose of these studies to analyze in detail the effect of human-likeness in a robot on the user's perception of and interaction with the robot.} There is not much research that investigates anthropomorphism based on real interaction between a human and a robot, and above all in a natural setting and over an extended period of time. 
Our data suggest that there are qualitative differences and dynamics over time in how participants relate to different types of robots. More concretely, we found that how human-like people perceived and interacted with the robot was not only based on how much the robot resembled a human (mostly in its behavioral cues)
but that this was determined by multiple factors, including individual
characteristics of the human user but also the context of use, for instance. Interestingly, our findings also suggest that anthropomorphism is likely to change over time with growing experience of the
user with the system.

Based on these observations and an extensive literature review, we would like to propose a model that accounts for variances and dynamics in anthropomorphism. This model of the ``Dynamics of Anthropomorphism'' considers variances in respect to multiple factors, including the human user, the anthropomorphic design of the non-human agent, the context of use, and the interaction experience (time).

Thereby, we would like to extend the current understanding of anthropomorphism in robotics and human-robot interaction.

@@

\subsection{A few results from the field}

\subsubsection{The Roomba experiment}

@@

\subsubsection{The Ranger experiment}

@@

\fxnote{state here our main research hypothesis}
\fxnote{state here our main contributions}
\fxnote{Introduce the remainder of the paper}

>>>>

%{Intro when start with anthropomorphic design}
The Uncanny Valley hypothesis \citep{mori_uncanny_1970} proposes that human-like designs of objects and non-human agents evoke a positive feeling of affinity in the human observer / user, which in turn reflects the experience of emotional engagement with and feelings of empathy for the non-human agent \citep{cheetham_human_2011}. The positive effect of so-called \textbf{anthropomorphic form} is exploited in the design of artificial agents (\eg digital characters, robots) for several reasons. Amongst others, anthropomorphic design should serve to evoke a positive valence as the Uncanny Valley hypothesis suggests, which in turn should foster people's acceptance of the artificial agent, and enable effective (human-like and especially human-social) interaction. 
However, at a certain point as the non-human agent approaches but fails to closely resemble a human, the evoked valence in the human user suddenly turns into a negative uncanny feeling of what \cite{mori_uncanny_1970} describes as disgust and revulsion. This, in turn, might lead to rejection of the non-human but human-like agent.
The Uncanny Valley hypothesis predicts that with further increasing the human-likeness of the non-human agent such that it appears so realistic that it is not distinguishable from a human but perceived to be human, the effect is again a positive one, reflecting feelings and engagement that are similar as those evoked by a human-human interaction.
Of course, roboticists and developers try to avoid ``falling'' into this Uncanny Valley with their human-like designs of artificial agents. Consequently, there is a considerable research effort into better understanding the boundaries of anthropomorphic design and their effect on the interaction and experience. However, findings have been inconsistent \citep{cheetham_human_2011}.

A phenomenon that is often studied related to the human-like design of robots is \textbf{anthropomorphism}. Besides a feeling of affinity, and empathy, anthropomorphism can be understood as being one of the resulting (and  more or less desired) effects of the anthropomorphic design of artificial agents (up to the first positive peak). The essence of anthropomorphism is the human user's perception of human-likeness in the non-human agent. Logically, anthropomorphism is said to be enhanced with increased human-like design of the agent. There are various other reasons and explanations for why people tend to anthropomorphize non-human agents. In general, as anthropomorphism emerges in a (real or imagined) interaction \citep{persson_anthropomorphism_2000} between at least two players (the human who anthropomorphizes and the non-human agent which is anthropomorphized) there are also two main perspectives on the phenomenon: an artifact-centered (design) and a human-centered (cognitive, psychological) one. The Uncanny Valley hypothesis can be categorized as being part of first, artifact-centered approach, to understand anthropomorphism.\\

We observed anthropomorphism in several short- and long-term human-robot interaction studies that we carried out with embodied robots, mostly in real-world settings. There is not much research that investigates anthropomorphism based on real interaction between a human and a robot, and above all in natural settings and over extended periods of time.

Our data suggest that there are qualitative differences and dynamics over time in how humans relate to different types of robots. More concretely, we found that people's perception of human-likeness in the robot and in turn their human-social engagement with the robot was not only based on how much the robot resembled a human (mostly in its behavioral cues)
but that this was determined by multiple factors, including individual
characteristics of the human user but also the context of use, for instance. Interestingly, our findings also suggest that how far people anthropomorphize a robot is likely to change over time with growing experience of the
user with the system.

Based on these observations and an extensive literature review, we would like to propose a model that accounts for variances and dynamics in anthropomorphism. This model of the \textit{``Dynamics of Anthropomorphism''} considers several impact factors, including the human user, the anthropomorphic design of the non-human agent, the context of use, and the interaction experience (time). With this initial model, we would like to extend the current understanding of anthropomorphism in robotics and human-robot interaction. This might in turn lead to a discussion about the short- and long-term benefits and other effects of the human-like (anthropomorphic) design of robots and other artificial agents.\\

The remainder of this article is organized as follows. 
First, in Section \ref{sec:overview-studies}, we give a brief overview on the various studies that we conducted to investigate anthropomorphism, namely (1) the Roomba study, (2) the Forum study, (3) the Ranger domino study. As an outcome of these studies, we propose two main research hypotheses related to anthropomorphism.
Next, in Section \ref{sec:anthropomorphism} we investigate these hypotheses by synthesizing literature on anthropomorphism. This should also serve to establish a common ground. We first explore an artifact-centered perspective and briefly outline how the human-like design of agents can encourage anthropomorphism. Then, we explore a human-centered perspective by integrating theories from cognitive science, developmental and social psychology that serve to explain anthropomorphism. At the end of this section we bring both perspectives together and interpret that anthropomorphism is a multi-factor phenomenon.
Based on the two previous sections, we propose an initial model of the \textit{``Dynamics of Anthropomorphism''} in Section \ref{sec:dynamics-model}. This model describes the variances and dynamics of people's tendency to anthropomorphize an artificial agent, such as a robot.
Next, in Section \ref{sec:cognition-neuroscience} we present a cognitive and neuroscience perspective on anthropomorphism to consolidate our considerations and the proposed model.
This will be followed by a discussion about the limitations of our approach (Section \ref{sec:discussion}). Further, we debate existing and alternative ways to measure
anthropomorphism in human-robot interactions.
Finally, Section \ref{sec:conclusion} concludes this article and offers starting points for further research on anthropomorphism and the effect of the human-like design approach in robotics.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		OVERVIEW ON OUR STUDIES
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{A few results from the field}
\label{sec:overview-studies}

In this section we give a brief overview of the studies that we conducted to investigate anthropomorphism. Note that it was not the main purpose of these studies to analyze in detail the effect of different dimensions human-likeness of a robot on the user's perception of and interaction with the robot. Each of these studies

The strong points

\subsection{Roomba study: how people live with a vacuum cleaning robot}

@@

\subsection{Forum study: using anthropomorphic language to describe a robot}

\hl{Julia: I think we should also include this study}

\subsection{Ranger domino study: effect of unexpected robot behavior}

@@

\fxnote{state here our main research hypothesis}
\fxnote{state here our main contributions}
